---
title: "Happiness and Alcohol Consumption"
output:
  github_document: default
  html_notebook: default
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(caret)
library(broom)
library(ggcorrplot)
library(stringr)
```


```{r}
data <- read.csv("data/HappinessAlcoholConsumption.csv")
str(data)
skimr::skim(data)

## Data Transformations
data_longer <- data %>% 
      pivot_longer(c(Beer_PerCapita:Wine_PerCapita), names_to = "Beverage", values_to = "Consumption_PerCapita") 

total_consumption <- data_longer %>%
  group_by(Country) %>%
  mutate(total_consumption = sum(Consumption_PerCapita))
```

## Boxplot of Happiness score against Region 
```{r}
ggplot(data, aes(x = Region, y = HappinessScore, fill = Region)) +
  geom_boxplot() +
  theme_classic() +
  labs(y = "Happiness Score", title = "Happiness score by Region (Boxplot)") +
  theme(
  legend.position = "none",
  axis.text.x = element_text(angle = 45)
  )


total_consumption %>%
  ggplot(aes(Region, total_consumption, fill = Region)) +
  geom_boxplot() +
  labs(y = "Total Consumption Per Capita", title = "Total Consumption of Alcohol Per Capita by Region") +
    theme_classic() +
  theme(
  legend.position = "none",
  axis.text.x = element_text(angle = 45)
  )
```

It is possible for countries and regions to have different cultures regarding alcohol consumption, as some cultures tend to drink more alcohol and others tend to drink less or completely abstain from drinking alcohol. Aside from the cultural difference between regions, a large impact of the overall will be the wealth, prosperity, health, and the living conditions of the region variables that are covered or correlated with the GDP per capita of the region/country as well as the Human Development Index (HDI).

## Boxplot of Happiness score against Region 
```{r}
data %>%
  mutate(Hemisphere = str_replace(Hemisphere, "noth", "north")) %>%
ggplot(aes(x = Hemisphere, y = HappinessScore)) +
  geom_boxplot() + 
  labs(x = "Hemisphere", y = "Happiness Score", title = "Happiness score by Hemisphere (Boxplot)") +
  theme_classic() +
  theme(
  legend.position = "none",
  axis.text.x = element_text(angle = 45)
  )
```
In the analysis, the consideration that there was a difference in happiness score based on the hemisphere a person lives in was not left out of the question because there have been significant studies around sunlight and the hormonal effects on an individual. So it is possible that those who are on or around the equator (people who tend to have more exposure to the sun) will be happier than those who are closer to the northern or southern pole of the Earth. However, based on the box plot presented above, it shows that there is very little difference in the overall happiness score of countries based on the hemisphere they reside in.


## Correlation of Data
```{r}
corr <- cor(data[,-c(1:3)])
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE,
   title = "Correlation Plot",
   colors = c("#6D9EC1", "white", "#E46726"))
```

Surprisingly, GDP per capita has a negative correlation with other variables such as HDI and Happiness score as common intuition would be that there is a positive relationship between GDP and HDI. This can be logically induced as an individual would in general be relatively happier with a higher wage and everything else remaining constant. A non-linear relationship should be considered between this variable and the others as a higher wage may only produce overall marginal effects at a certain point.


## Relationship between happiness score and Human development index (HDI)
```{r}
#Exploring the happiness score against human development index
## Due to possible strong linear correlation with happiness

ggplot(data, aes(HDI, HappinessScore, color = Region)) +
  geom_point() +
  labs(x = "Human Development Index", y = "Happiness Score", title = "Happiness Score against Human Development Index") +
  theme_classic() +
  geom_abline(intercept = coef(lm(HappinessScore ~ HDI, data = data))[1], 
              slope = coef(lm(HappinessScore ~ HDI, data = data))[2],
              color = "lightblue",
              size = 1)

```

The first variable I chose to explore is HDI due to the fact that it has the strongest linear correlation with happiness score and should be included into any machine learning model for this reason. In addition, visualizing it allows for analysis of different possible patterns and to view the overall dispersion of the data against a regression.

## Exploring GDP per Capita
```{r}
#GDP per capita against Happiness Score
## Exploring GDP against Happiness score due to counter-intuitive results

ggplot(data, aes(GDP_PerCapita, HappinessScore, color = Region))+
  geom_point() +
  theme_classic() +
  labs(x = "GDP Per Capita", y = "Happiness Score", title = "Happiness Score against GDP Per Capita", caption = "*Countries with sub 1k (USD) GDP Per Capita are standardized differently")

ggplot(data, aes(GDP_PerCapita))+
  geom_histogram(fill = "lightblue") +
  theme_classic()
```

Based on the above graphs, an error in addition to a linear transformation needs to take place to deal with the variable. There is an interesting thing occurring where a large portion of countries with a GDP per capita of greater than 250 have some of the lowest happiness scores. But they do not exceed a GDP of 1000: moreover, the current country with the lowest GDP per Capita is Zimbabwe at 1.029, thus, I am lead to the conclusion that there has been an error in the transformation of this variable and countries with a were not properly transformed to GDP per Capita in thousands.

### Corrected GDP per Capita
```{r}
# Correcting GDP
data <- data %>%
  mutate(GDP_PerCapita = ifelse(GDP_PerCapita > 250, GDP_PerCapita/1000, GDP_PerCapita))

ggplot(data, aes(GDP_PerCapita, HappinessScore, color = Region))+
  geom_point() +
  theme_classic() +
  labs(x = "GDP Per Capita", y = "Happiness Score", title = "Happiness Score against GDP Per Capita")

ggplot(data, aes(GDP_PerCapita))+
  geom_histogram(fill = "lightblue") +
  theme_classic() + 
  labs(x = "GDP Per Capita (Log Transformation)", y = "Count", title = "Histogram of GDP Per Capita")
```

After correcting for this error, it can been seen that there is a curvilinear relationship between GDP per capita and happiness score, which is easily address with a log transformation of GDP per capita. Also, the graph shows more dispersion with countries that have GDP per capita between 10 and 60.

```{r}
# Correcting GDP
data <- data %>%
  mutate(GDP_PerCapita_log = log(GDP_PerCapita))

ggplot(data, aes(GDP_PerCapita_log, HappinessScore, color = Region))+
  geom_point() +
  theme_classic() +
  labs(x = "GDP Per Capita Log Transformation", y = "Happiness Score", title = "Happiness Score against GDP Per Capita (Log Transformation)")

ggplot(data, aes(GDP_PerCapita_log))+
  geom_histogram(aes(y = ..density..), fill = "lightblue") +
  theme_classic() +
  geom_density() +
  labs(x = "GDP Per Capita", y = "Probability", title = "Histogram of GDP Per Capita")

corr2 <- cor(data[,-c(1:3)])
ggcorrplot(corr2, hc.order = TRUE, type = "lower",
           lab = TRUE,
           title = "Correlation Plot",
   colors = c("#6D9EC1", "white", "#E46726"))
```

In the above graphs, GDP per capita which had a negative correlation coefficient is now strongly positive with correlated with happiness score after the correction in the data and the log transformation.   


## Exploring Alcohol and Happiness score
```{r}
data_longer%>%
ggplot(aes(Consumption_PerCapita, HappinessScore, color = Beverage)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +
  labs(x = "Consumption of Alcohol", y = "Happiness Score", title = "Happiness score against Consumption of Alcohol by Beverage Type")

```

## Consumption Based on Region
```{r}
# linear Regression
mod1 <- lm(HappinessScore ~ Region + GDP_PerCapita_log + HDI + Beer_PerCapita + Spirit_PerCapita + Wine_PerCapita, data = data)
mod2 <- lm(HappinessScore ~ Region + GDP_PerCapita_log + HDI + log(Beer_PerCapita) + log(Spirit_PerCapita) + log(Wine_PerCapita), data = data)
mod3 <- lm(HappinessScore ~ GDP_PerCapita_log + HDI + log(Beer_PerCapita) + log(Spirit_PerCapita) + log(Wine_PerCapita), data = data)
mod4 <- lm(HappinessScore ~ log(Beer_PerCapita) + log(Spirit_PerCapita) + log(Wine_PerCapita), data = data)

mod1 %>% summary()
mod2 %>% summary()
mod4 %>% summary()

Models <- tibble(model = c("mod1", "mod2", "mod3"), coeffs = list(tidy(mod1), tidy(mod2), tidy(mod3)), modfit = list(glance(mod1), glance(mod2), glance(mod3)))
```
When looking at the models they appear to have a strong fit, however, it appears that a lot of the variance captured by GDP and HDI and region are also in the variables Beer_perCapita, Spirit_PerCapita, and Wine_PerCapita. The coefficient of variation is `r .503/mean(data$HappinessScore)`, and has a Residual standard error of 0.503 meaning that on average the model estimates within 0.503 happiness score, and the overall model has a coefficient of determination R-squared of 0.8289, meaning that it explains 82.89% of the variation in the data.

## Conclusion
For this project, I quickly analyzed the data using regression methods, due to the shape and style of the data, it seemed like a solid model that would be extraordinarily useful in this situation. The overall
